%\enableregime[utf]  % use UTF-8

\setupcolors[state=start]
\setupinteraction[state=start, color=blue] % needed for hyperlinks

\usemodule[simplefonts]
\setmainfont[Liberation-Serif]
\setmonofont[inconsolata]

\setuppapersize[A4][A4]  % use letter paper
%\setuplayout[width=middle, backspace=1.5in, cutspace=1.5in,
%             height=middle, header=0.75in, footer=0.75in] % page layout
\setuppagenumbering[location={footer,right}]  % number pages
\setupbodyfont[12pt]  % 11pt font
\setupwhitespace[small]  % inter-paragraph spacing

\setupindenting[medium]
\indenting[always]

\setuphead[section][style=\tfc]
\setuphead[subsection][style=\tfb]
\setuphead[subsubsection][style=\bf]

% define descr (for definition lists)
\definedescription[descr][
  headstyle=bold,style=normal,align=left,location=hanging,
  width=broad,margin=1cm]

% prevent orphaned list intros
\setupitemize[autointro]

% define defaults for bulleted lists 
\setupitemize[1][symbol=1][indentnext=no]
\setupitemize[2][symbol=2][indentnext=no]
\setupitemize[3][symbol=3][indentnext=no]
\setupitemize[4][symbol=4][indentnext=no]

\setupthinrules[width=15em]  % width of horizontal rules

% for block quotations
\unprotect

\startvariables all
blockquote: blockquote
\stopvariables

\definedelimitedtext
[\v!blockquote][\v!quotation]

\setupdelimitedtext
[\v!blockquote]
[\c!left=,
\c!right=,
before={\blank[medium]},
after={\blank[medium]},
]

\protect

\starttext

\subject{\quotation{Media Hot and Cold}}

\subject{Mcluhan, Marshall}

\startitemize
\item
  \quotation{A hot medium is on that extends one single sense in \quote{high definition}. High definition is the state of being well filled with data.}
  \startitemize
  \item
    photograph: high def
  \item
    comics: low def (use Scott McCloud's argument on fidelity vs
    connection with comic characters)
  \item
    speeh: low def
    \quotation{because so little is given and so much has to be filled in by the listener}
  \stopitemize
\item
  because cool media leave much up to the audience, they can be seen
  as \quotation{high in participation or completion by the audience.}
  This is in contrast to hot media, which are low in participation
  precisely because it leaves so little for the audience to
  accomplish.
\item
  hieroglyphic/ideogrammic (cool) vs phonetic (hot) writing
\item
  \quotation{The alphabet, when pushed to a high degree of visual intensity, became typography.}
\item
  \quotation{The hotting-up of the medium of writing to repeatableprint intensity led to nationalism and the religious wars of the sixteenth century.}
\item
  \quotation{paper is a hot medium that serves to unify spaces horizontally, both in political and entertainment empires}
  \startitemize
  \item
    so, a return to the spatial rather than the temporal; is this the
    reason for European agnosticism? does simply reading detach you
    from the temporally-bound authority of the bible?
  \item
    even more importantly, if print serves this purpose, electronic
    screens amplifes the effect significantly
  \stopitemize
\item
  \quotation{With print many earlier forms were excluded from life and art, and many were given strange new intensity.}
\item
  \quotation{A tribal and feudal hierarchy of traditional kind collapses quickly when it meets any hot medium of the mechanical, uniform, and repetitive kind. The medium of money or wheel or writing, or any other form of specialist speed-up of exchange and information, will serve to fragment a tribal structure.}
\item
  \quotation{Specialist technologies detribalize. The nonspecialist electric technology retribalizes.}
  \startitemize
  \item
    don't specialist technologies also tribalize in that specialists
    form cliques and even adversarial relationships with other groups
    of specialiasts?
  \item
    \quotation{The process of upset resulting from a new distribution of skills is accompanied by much culture lag in which people feel compelled to look at new situations as if they were old ones\ldots{}}
  \stopitemize
\item
  the concentric spiral form, present in
  \quotation{the Hebrew and Eastern mode of thought} as defining of
  the electric age:
  \quotation{It is a redundant form inevitable to the electric age, in which the concentric pattern is imposed by the instant quality, and overlay in depth, of electric speed.}
\item
  \quotation{More and more we turn from the content of messages to study total effect.}
  \startitemize
  \item
    \quotation{Concern with effect rather than meaning is a basic change of our electric time, for effect involves the total situation, and not a single level information movement.}
  \stopitemize
\item
  \quotation{The effect of hot media treatment cannot include much empathy or participation at any time.}
\item
  \quotation{The price of eternal vigilance is indifference.}
\stopitemize

\stoptext

