% Links
\setupcolors[state=start]
\setupinteraction[state=start,color=darkblue]

% ..
\setuppagenumbering[location={footer,left,margin}]  

% paper, layout, etc.
\setuppapersize[A4]
%\setuplayout[width=6.5in,height=10.5in,topspace=0.5in,backspace=1in,
%  header=0.5in,footer=0.5in]
% double space
%\setupinterlinespace[line=5.6ex]
% 1/2 inch indents
\setupindenting[medium]
\indenting[always]
% CSL is a new form of handling bibliographic data and citations. It also apparently   http://xbiblio.sourceforge.net/csl/
% jagged-right (aligned to the left)
\setupalign[right]
% font
%\setupbodyfont[rm,12pt]
\usemodule[simplefonts]
\setmainfont[Liberation-Serif]
\setmonofont[inconsolata]

\setupwhitespace[medium]

% for long quotes
\definestartstop[longquote][
  before={\indenting[never]
    \setupnarrower[left=0.5in,right=0.5in]
    \startnarrower[2*left,right]},
  after={\stopnarrower
    \indenting[yes]}]

% for heading and header
\def\MLA#1[#2][#3][#4][#5][#6][#7]%
	{\setuppagenumbering[left=#3 ,location={header,right}]
	\indenting[never]
	#2 #3\par#4\par#5\par#6\par\startalignment[middle]#7\stopalignment
	\indenting[yes]}
	
% following hanging indent code (also in workscited) taken from 
%  http://www.ntg.nl/pipermail/ntg-context/2004/005280.html
% [NTG-context] Re: Again: "hanging" for a lot of paragraphs?
%  ~ Patrick Gundlach
\def\hangover{\hangafter=1\hangindent=0.5in}
\definestartstop[workscited][
  before={
    \page[no]
    \indenting[never]
    \startalignment[left]
    \subject{Literature}
    \stopalignment
    \bgroup\appendtoks\hangover\to\everypar
    },
  after={
    \egroup
    \indenting[yes]}]


\starttext      


\starttabulate[|l|l|]
\NC Student:
\NC John Haltiwanger
\NR
\NC Paper:
\NC Literature list
\NR
\NC Supervisor:
\NC Richard Rogers
\NR
\NC Date:
\NC 6 March 2010
\NR 
\stoptabulate  

\startworkscited

Bolter, Jay David.
{\em Writing Space: The computer, hypertext, and the history of writing}.
Hillsdale, New Jersey: Lawrence Erlbaum, 1991. Print.

Bolter, Jay David and Richard Grusin. {\em Remediation}. Cambridge:
MIT Press, 1999. Print.

Cramer, Florian.
{\em Words Made Flesh: Code, culture, imagination}. Piet Zwart
Institute, 2005. Web.
\letterless{}\useURL[1][http://pzwart.wdka.hro.nl/mdr/research/fcramer/wordsmadeflesh/][][http://pzwart.wdka.hro.nl/mdr/research/fcramer/wordsmadeflesh/]\from[1]\lettermore{}

Dworkin, Craig. \quotation{Hypermnesia}. {\em boundary 2}, 36:3, p.
77--95. 2009.

Fuller, Matthew.
\quotation{It looks like you're trying to write a letter: Microsoft Word}.
2000. Web.
\letterless{}\useURL[2][http://www.nettime.org/Lists-Archives/nettime-l-0009/msg00040.html][][http://www.nettime.org/Lists-Archives/nettime-l--0009/msg00040.html]\from[2]\lettermore{}

Fuller, Matthew, ed. {\em Software studies: a lexicon}. Cambridge:
MIT Press, 2008. Print.

Gitelman, Lisa.
{\em Always Already New: Media, history, and the data of culture}.
Cambridge: MIT Press, 2008. Print.

Hagen, Hans. {\em ConTeXt: A brief excursion}. Pragma ADE, 1999.
Web.
\letterless{}\useURL[3][http://www.pragma-ade.com/general/manuals/ms-cb-en.pdf][][http://www.pragma-ade.com/general/manuals/ms-cb-en.pdf]\from[3]\lettermore{}

Hagen, Hans. {\em ConTeXt: the manual}. Pragma ADE, 2001. Web.
\letterless{}\useURL[4][http://www.pragma-ade.com/general/manuals/cont-eni.pdf][][http://www.pragma-ade.com/general/manuals/cont-eni.pdf]\from[4]\lettermore{}

Hagen, Hans. {\em The history of luaTeX}. Pragma ADE, 2009. Web.
\letterless{}\useURL[5][http://www.pragma-ade.com/general/manuals/mk.pdf][][http://www.pragma-ade.com/general/manuals/mk.pdf]\from[5]\lettermore{}

Hayles, N. Katherine.
{\em Electronic Literature: New horizons for the literary}.
Indiana: University of Notre Dame, 2008. Print.

Heim, Michael.
{\em Electric Language: A Philosophical Study of Word Processing}.
New Haven: Yale, 1989. Print.

Illich, Ivan and Barry Sanders.
{\em ABC: The alphabetization of the popular mind}. San Francisco:
North Point Press, 1988. Print.

Kinross, Robin. {\em Modern typography}. London: Hyphen Press,
2004. Print.

Knuth, Donald. {\em The TeXBook}. Massachusetts: Addison-Wesley,
1984. Print.

Knuth, Donald. {\em Digital Typography}. Stanford: CSLI, 1999.

Latour, Bruno. {\em Reassembling the social}. Oxford: Oxford
University Press, 2005. Print.

Manovich, Lev. Software Takes Command. Forthcoming from The MIT
Press in 2010 in Software Studies Series. Web.
\letterless{}\useURL[6][http://softwarestudies.com/softbook/manovich_softbook_11_20_2008.pdf][][http://softwarestudies.com/softbook/manovich\letterunderscore{}softbook\letterunderscore{}11\letterunderscore{}20\letterunderscore{}2008.pdf]\from[6]\lettermore{}

McLuhan, Marshall and Wilfred Watson.
{\em From Cliche to Archetype}. New York: Viking Press, 1970.
Print.

\stopworkscited

\stoptext