%\enableregime[utf]  % use UTF-8

\setupcolors[state=start]
\setupinteraction[state=start, color=blue] % needed for hyperlinks

\usemodule[simplefonts]
\setmainfont[Liberation-Serif]
\setmonofont[inconsolata]

\setuppapersize[A4][A4]  % use letter paper
%\setuplayout[width=middle, backspace=1.5in, cutspace=1.5in,
%             height=middle, header=0.75in, footer=0.75in] % page layout
\setuppagenumbering[location={footer,right}]  % number pages
\setupbodyfont[12pt]  % 11pt font
\setupwhitespace[small]  % inter-paragraph spacing

\setupindenting[medium]
\indenting[always]

\setuphead[section][style=\tfc]
\setuphead[subsection][style=\tfb]
\setuphead[subsubsection][style=\bf]

% define descr (for definition lists)
\definedescription[descr][
  headstyle=bold,style=normal,align=left,location=hanging,
  width=broad,margin=1cm]

% prevent orphaned list intros
\setupitemize[autointro]

% define defaults for bulleted lists 
\setupitemize[1][symbol=1][indentnext=no]
\setupitemize[2][symbol=2][indentnext=no]
\setupitemize[3][symbol=3][indentnext=no]
\setupitemize[4][symbol=4][indentnext=no]

\setupthinrules[width=15em]  % width of horizontal rules

% for block quotations
\unprotect

\startvariables all
blockquote: blockquote
\stopvariables

\definedelimitedtext
[\v!blockquote][\v!quotation]

\setupdelimitedtext
[\v!blockquote]
[\c!left=,
\c!right=,
before={\blank[medium]},
after={\blank[medium]},
]

\protect

\starttext

\subject{Bibliography}

Bolter, Jay David. (2001).
{\em Writing Space: Computers, hypertext, and the remediation of print}.
New York: Routledge. Print.

Bolter, Jay David and Richard A. Grusin. (1996).
\quotation{Remediation}. {\em Configurations} 4:3. PDF.

Bringhurst, Robert. (2008).
{\em The Elements of Typographic Style, version 3.2}. Vancouver:
Hartley \& Marks. Print.

Brooks, Frederick. (1975).
{\em The Mythical Man-Month: Essays on software engineering}.
Massachusetts: Addison-Wesley. PDF.

Cramer, Florian. (2001).
\quotation{Digital Code and Literary Text}. {\em netzlituratur}.
Web.
\letterless{}\useURL[1][http://www.netzliteratur.net/cramer/digital_code_and_literary_text.html][][http://www.netzliteratur.net/cramer/digital\letterunderscore{}code\letterunderscore{}and\letterunderscore{}literary\letterunderscore{}text.html]\from[1]\lettermore{}
(last accessed 5 June 2010).

Cramer, Florian and Matthew Fuller. (2008). \quotation{Interface.}
In {\em Software Studies: a lexicon}, edited by Matthew Fuller. MIT
Press: Cambridge.

Fuller, Matthew.
\quotation{It looks like you're trying to write a letter: Microsoft Word}.
2000. Web.
\letterless{}\useURL[2][http://www.nettime.org/Lists-Archives/nettime-l-0009/msg00040.html][][http://www.nettime.org/Lists-Archives/nettime-l--0009/msg00040.html]\from[2]\lettermore{}
(last accessed 5 June 2010).

Galloway, Alexander. (2010). \quotation{Interface}. Presented at
{\em A wedge between public and private conference} on 22 April
2010, Amsterdam.

Gitelman, Lisa. (2008).
{\em Always Already New: Media, history, and the data of culture}.
Cambridge: MIT Press. Print.

Hagen, Hans. (2009). {\em The history of luaTeX}. Netherlands:
Pragma ADE. Web.
\letterless{}\useURL[3][http://www.pragma-ade.com/general/manuals/mk.pdf][][http://www.pragma-ade.com/general/manuals/mk.pdf]\from[3]\lettermore{}
(last accessed 5 June 2010).

Hayles, N. Katherine. (2004)
\quotation{Print is Flat, Code is Deep: The Importance of Media Specific Analysis}.
{\em Poetics Today} 25:1. PDF.

Hayles, N. Katherine. (2008).
{\em Electronic Literature: new horizons for the literary}. Notre
Dame: University of Notre Dame. Print.

Kinross, Robin. (2004).
{\em Modern typography: an essay in critical history}. London:
Hyphen Press. Print.

Knuth, Donald. (1984). {\em The TeXBook}. Massachusetts:
Addison-Wesley. Print.

Knuth, Donald. (1999). {\em Digital Typography}. Stanford: CSLI.
Print.

Manovich, Lev. (2001). {\em The Language of New Media}. MIT Press:
Cambridge. Print.

Rosenberg, Scott. (2008). {\em Dreaming in Code}. New York: Three
Rivers Press. Print.

Simondon, Gilbert. (2009).
\quotation{The Position of the Problem of Ontogenesis}. Gregor
Flanders, trans. {\em Parrhesia} 7. Web.
\letterless{}\useURL[4][http://www.parrhesiajournal.org/parrhesia07/parrhesia07_simondon1.pdf][][http://www.parrhesiajournal.org/parrhesia07/parrhesia07\letterunderscore{}simondon1.pdf]\from[4]\lettermore{}
(last accessed 5 June 2010).

Snelting, Femke. (2009). \quotation{The Making of}.
{\em Tracks in electr(on)ic fields}. Brussels: Constant. Web.
\useURL[5][http://ospublish.constantvzw.org/wp-content/uploads/makingof.pdf][][http://ospublish.constantvzw.org/wp-content/uploads/makingof.pdf]\from[5]
(last accessed 5 June 2010).

\stoptext

